### 2023-08-21 thesis/sections/2-background.tex
review
introduction, ethical considerations

\cmtDK[inline]{7 pages}

\subsection{Referring expressions}
Even though, recent large language models (LLMs) such as BERT (\citet{Devlin2018}) or GPT-3 (\citet{Brown2020}) produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
In this way, their only way of learning, how to generate language as well as solve these tasks is to find patterns of how words and sentences are used in these large corpora.
Many researchers criticize this approach, by arguing that these model can't learn meaning just by learning on text.
In \citet{Bender2020}, the authors argue that meaning is bound to a communicative intent.
These are the purposes of, why humans are using language.
This intent is connected to the real world that exists outside of language, in other words it is grounded in the world.
This grounding is missing, when models only train on abstract textual representations of the world.
\citet{Bisk2020} argues that a multimodal approach, including for instance perception as well as social context, is needed to learn meaning in a broader context.
One added modality to ground language is often vision. \cmtDK{cite}
Hereby, the model needs to learn how to associate linguistic concepts in text corpora with features, extracted from visual input.
For instance, a model can learn to associate the noun "dog" with an animal seen in an image or associate the action of "jumping over" with the animal being above an object.

@@

The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge (\citet{Kirby2002}).
Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.

>>>>

\subsection{Referring expressions}
Even though, recent large language models (LLMs) such as BERT \citep{Devlin2019} or GPT-3 \citep{Brown2020} produce impressive results on many tasks in Natural Language Processing, they are often only trained on big amounts of text corpora.
In this way, their only way of learning, how to generate language as well as solve these tasks is to find patterns of how words and sentences are used in these large corpora.
Many researchers criticize this approach, by arguing that these model can't learn meaning just by learning on text.
In \citet{Bender2020}, the authors argue that meaning is bound to a communicative intent.
These are the purposes of, why humans are using language.
This intent is connected to the real world that exists outside of language, in other words it is grounded in the world.
This grounding is missing, when models only train on abstract textual representations of the world.
\citet{Bisk2020} argues that a multimodal approach, including for instance perception as well as social context, is needed to learn meaning in a broader context.
One added modality to ground language is often vision. \cmtDK{cite}
Hereby, the model needs to learn how to associate linguistic concepts in text corpora with features, extracted from visual input.
For instance, a model can learn to associate the noun "dog" with an animal seen in an image or associate the action of "jumping over" with the animal being above an object.

@@

The entities can be set into real or simulated interactive situation, a language game, that meaning and a language can emerge \citep{Kirby2002}.
Multiple deep neural models, called agents, communicate with a set of symbols to solve a task.
Guided by the task and the rules, how agents can generate and interpret symbols, a language can emerge.

### 2023-08-21 thesis/sections/3-methodology.tex
introduction, ethical considerations

\subsection{CLEVR dataset}
The basis for my datasets is the Visual Question Answering (VQA) dataset CLEVR \citep{Johnson2016}.
This dataset is split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
Many of the existing VQA datasets come with two problems.
First, they include many biases, such as biases in the base images and biases in the linguistic properties of the questions and answers.
A relatively high number of images of dogs in a dataset, might for instance bias a classifier model towards classifying dogs most of the time.
On the other hand, repeating patterns in the questions and answers might also be exploited by a model, without extracting the needed information from the image.
For these reasons the CLEVR dataset aims to reduce the biases as much as possible in both images and questions and answers.
Secondly, datasets may come with only a limited amount of annotations and information about the state in an image.
The CLEVR dataset uses artificially rendered 3D-scenes.
By doing so, all information about for instance the location of objects or their relations to each other can be saved and later used in training models or analyzing their results.

@@

\subsection{Feature extractors}
In computer vision tasks, machines need to analyze images and extract information from them.
To do this, machines often rely on feature extractors.
Features are important parts or patterns in an image, which can have different levels of abstractions.
They can for example be low-level features, as geometric information about lines and edges in an image, or also very abstract information about whole objects.
Traditional approaches involve extracting key points and descriptors from an image and using them to represent the image (\cite{Harris1988}, \cite{Lowe1999}, \cite{Bay2006}).
More recently, convolutional neural networks (CNN) became popular due to their ability to learn complex features automatically from raw image data.
These are also used in this thesis as a first layer to extract important information from the image. Hereby, two different architectures are tested.

@@

\subsection{Ethical considerations}

>>>>

\subsection{CLEVR dataset}
The basis for my datasets is the Visual Question Answering (VQA) dataset CLEVR \citep{Johnson2017a}.
This dataset is split in two parts: a collection of images, and a set of questions and answers that refer and describe each image.
Many of the existing VQA datasets come with two problems.
First, they include many biases, such as biases in the base images and biases in the linguistic properties of the questions and answers.
A relatively high number of images of dogs in a dataset, might for instance bias a classifier model towards classifying dogs most of the time.
On the other hand, repeating patterns in the questions and answers might also be exploited by a model, without extracting the needed information from the image.
For these reasons the CLEVR dataset aims to reduce the biases as much as possible in both images and questions and answers.
Secondly, datasets may come with only a limited amount of annotations and information about the state in an image.
The CLEVR dataset uses artificially rendered 3D-scenes.
By doing so, all information about for instance the location of objects or their relations to each other can be saved and later used in training models or analyzing their results.

@@

\subsection{Feature extractors}
In computer vision tasks, machines need to analyze images and extract information from them.
To do this, machines often rely on feature extractors.
Features are important parts or patterns in an image, which can have different levels of abstractions.
They can for example be low-level features, as geometric information about lines and edges in an image, or also very abstract information about whole objects.
Traditional approaches involve extracting key points and descriptors from an image and using them to represent the image \citep{Harris1988,Lowe1999,Bay2006}.
More recently, convolutional neural networks (CNN) became popular due to their ability to learn complex features automatically from raw image data.
These are also used in this thesis as a first layer to extract important information from the image. Hereby, two different architectures are tested.

@@

\subsection{Ethical considerations}
In the field of natural language processing (NLP) ethical issues often play major roles.
These can be part of the used datasets, the created models and their training as well as the application of the models.
For datasets, the role data privacy is increasing with the necessity of larger amounts of data \citep{Klymenko2022}.
Furthermore, datasets often contain biases, based for instance on the authors of the collected natural language texts. Often they also contain biases such as overrepresentations and underrepresentations.
Even though some of these biases are inherent to the data and not necessarily negative, much research is indicating that undetected and unaddressed biases in datasets might lead to negative consequences \citep{Shah2020,Field2021,Bender2021}.
Training neural models can also lead to environmental issues, as large models need to process huge amounts of data and require a lot of energy \citep{Bender2021}.
Finally, the application of trained models can create harm.
This applies for example to easy accessible large language model (LLMs) that can be used to create information hazard \citep{Weidinger2022}.

The research in this thesis tries to reduce these risks.
Looking at the datasets that are used in this thesis, all data is created artificially and contains therefore no personal information.
Even further, the aim of the creation of these datasets is to reduce biases.
It doesn't include any social information, but on the other hand consists only of abstract scenes.
The choice of which attributes the objects are made up is inherently biased towards human cognition, but doesn't have a social impact.
The models and agents are therefore trained, by including as few human biases as possible.

Looking at the environmental issues, the models used in this thesis consist of only few trained layers and the training is therefore short and doesn't require much energy.
Larger models as the feature extractors are already pretrained and add no additional consumption.

Finally, the purpose of this thesis is to analyze the results and the emerged language and draw conclusions, how emerged languages can be grounded better in the environment.
For that reason, the final models can't be uses in any real world applications and produce potential harm.

### 2023-08-21 thesis/sections/4-exp-setup.tex
introduction, ethical considerations

\section{Experiments}
\label{sec:exp-setup}

@@

The experiments that are executed using language games have a similar structure as the pre-experiments, since those lied the basis for the language games.

@@

First, I will discuss \emph{discrimination games}, because they have the simplest setup.
Furthermore, other language games that research this topic use a very similar setup.
In the next step, I will look into caption generators that are set up as a language game.
Here, the sender describes the scene, while the receiver needs to generate a caption.
In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.

\cmtDK{look at emerged language: distribution of vocab/ngrams, how do they relate to the images (clustering?)}

@@

The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2016}.
The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
One central difference is the production of the message.
The main goal of their language game was the identification of the concept that and image was related to.
Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
It is therefore assumed that the sender will communicate these discriminative attributes.
For that reason, the sender is allowed to generate sequences as a message.

Both 'Dale' datasets are used to train the agents in this mode.
Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
As in \citet{Lazaridou2016}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
The sender is assumed to learn which of the input images is the target images by this approach.
The receiver receives the images in completely shuffled order.

@@

\cmtDK{look at language}
\cmtDK{similarity to bounding box classifier}

@@

To compare the emerged language to English in a quantitative way, a probing approach is used.
Probing can be used to analyze and interpret the hidden representations in neural networks.
Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \cite{Dale1995}.
Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
The encoder LSTM encodes the messages of the emerged language.
The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
The resulting vector is used as the initial state of the decoder LSTM.
The decoder LSTM is then trained to produce the English referring expression.
While decoding, teacher forcing is applied.
The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.

@@

Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
This fact concludes that all emerged languages don't rely on the GRE algorithm by \cite{Dale1995}.
Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.

>>>>

\section{Experimental setup and results}
\label{sec:exp-setup}

@@

The experiments that are executed using language games have a similar structure as the experiments in the previous chapters, since those provided the basis for the language games.

@@

First, I will discuss \emph{discrimination games}, because they have the simplest setup.
Furthermore, other language games that research this topic use a very similar setup.
In the next step, I will look into \emph{caption generators} that are set up as a language game.
Here, the sender describes the scene, while the receiver needs to generate a caption.
In the last step, I try to lose as much human bias as possible and the models are trained on just 'pointing' towards the target object, by again predicting its center coordinates.

The languages that emerge are analyzed in two ways.
An analysis of the frequency of used symbols and messages, as well as an examination for which images and objects they are used indicates the structure of the language and meanings of the symbols.
In a second step, the language is compared to several referring expressions in natural language.
By doing this, it can be seen if the models learned to use similar ways of referring to objects as humans do, or if they rely on different approaches.

@@

The discrimination games in this research have a very similar setup as described in \citet{Lazaridou2017}.
The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
One central difference is the production of the message.
The main goal of their language game was the identification of the concept that and image was related to.
Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
It is therefore assumed that the sender will communicate these discriminative attributes.
For that reason, the sender is allowed to generate sequences as a message.

Both 'Dale' datasets are used to train the agents in this mode.
Similarly to the bounding box classifier, bounding boxes around each of the objects are extracted and fed to the game.
As in \citet{Lazaridou2017}, the sender receives the bounding box of the target object as the first image, while the rest of the bounding boxes are shuffled.
The sender is assumed to learn which of the input images is the target images by this approach.
The receiver receives the images in completely shuffled order.

@@

\cmtDK{similarity to bounding box classifier}

@@

To compare the emerged language to English in a quantitative way, a probing approach is used.
Probing can be used to analyze and interpret the hidden representations in neural networks.
Hereby, a second neural model is trained to predict selected linguistic properties on the basis of the hidden representations.
If this model is successfully able to predict the linguistic properties, the hidden representations are connected to them.
If second model can't be trained, it indicates that there is no correlation between the hidden representation and the linguistic property.
In the case of the emerged language, a neural model is trained to translate the messages of the sender into English referring expressions, based on the GRE algorithm by \citet{Dale1995}.
Hereby, the model consists of an encoder LSTM, one linear layer and a decoder LSTM.
The encoder LSTM encodes the messages of the emerged language.
The final hidden state is used as the meaning of the complete sentence and passed through the linear layer, which should learn an abstract representation of the message.
The resulting vector is used as the initial state of the decoder LSTM.
The decoder LSTM is then trained to produce the English referring expression.
While decoding, teacher forcing is applied.
The success of the model is validated calculating cross entropy between the models predictions and the target English referring expression.
Since generalization doesn't play a role for probing, no dropout is applied and the model is trained and validated on the complete dataset; no test or validation split is used.

@@

Table \ref{tab:probing_discrimnator_language} shows the results for all different languages.
First, it can be seen for all emerged languages, the model is able to find correlations between natural language referring expressions and the messages by the sender.
Still, all losses stay high and closer to the loss of the baseline, namely where the model only learns the patterns in the English referring expressions.
This fact concludes that all emerged languages don't rely on the GRE algorithm by \citet{Dale1995}.
Nonetheless, it may be possible that a different algorithm is used to create referring expressions in the artificial language.
